---
title: "Homework 1"
author: Jeanette Shekelle
date: 09/21/18
output: github_document
---

## Loading tidyverse package
```{r}
library(tidyverse)
```

## Problem 1
Create a data frame comprised of:

- A random sample of size 10 from a uniform[0, 5] distribution.
- A logical vector indicating whether elements of the sample are greater than 2.
- A (length-10) character vector.
- A (length-10) factor vector.

```{r}
set.seed(1234)
problem_one_df = tibble(
  vec_numeric = runif(10, 0, 5),
  vec_char = c("elephant", "giraffe", "dog", "cat", "cow", "fish", "kangaroo", "zebra", "monkey", "snake"),
  vec_factor = factor(c("red", "orange", "yellow", "green", "blue", "purple", "pink", "white", "black", "grey")),
  vec_logic = (vec_numeric > 2))

```

## Taking the mean of each variable:

The numeric mean is `r mean(problem_one_df$vec_numeric)`.
The character mean is `r mean(problem_one_df$vec_char)`.
The factor mean is `r mean(problem_one_df$vec_factor)`.
The logic mean is `r mean(problem_one_df$vec_logic)`.

The charactor and factor means were NA because they are not numeric values. You can't take the mean of non-numeric variables. There is no ordering or numbering or charactors or factors. The logic variable's (true/false) mean is the proportion of trues to falses with trues being represented by 1 and falses being represented by 0. The numeric mean works because it is the mean of the numbers from 0 to 5.  

## Applying the as.numeric function to the logical, character, and factor variables:

```{r eval = FALSE} 
as.numeric(problem_one_df$vec_char)
as.numeric(problem_one_df$vec_factor)
as.numeric(problem_one_df$vec_logic)
```

The character variable was not converted to a numeric value because there was no pattern or order to turn into numbers. The conversion fails. The factor variable was converted into random numbers 1-10, with all different numbers because each factor is a different color, and the logic vector was converted into 0s for false and 1s for true. 

Converting the character variable to a factor variable to a numeric variable:
```{r eval = FALSE}
as.numeric(as.factor(problem_one_df$vec_char))
```
The character variable is assigned levels when converted to a factor variable. The levels can then be converted to numbers (1-10).


Converting the factor variable to a character variable to a numeric variable:
```{r eval = FALSE}
as.numeric(as.character(problem_one_df$vec_factor))

```
When the factor variable is converted to a character variable, it is assigned letters. Letters cannot be turned into numbers so the conversion fails. We see a warning message and all the values are turned to NA. 

## Problem 2
Create a data frame comprised of:

- x: a random sample of size 1000 from a standard Normal distribution
- y: a random sample of size 1000 from a standard Normal distribution
- A logical vector indicating whether the x + y > 0
- A numeric vector created by coercing the above logical vector
- A factor vector created by coercing the above logical vector

```{r}

problem_two_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  vec_logic = (x + y > 0)
)

problem_two_df$vec_numeric = as.numeric(problem_two_df$vec_logic)
problem_two_df$vec_factor = as.factor(problem_two_df$vec_logic)
```

The size of the dataset is  `r nrow(problem_two_df)`.
The mean of x is `r mean(problem_two_df$x)`.
The median of x is `r median(problem_two_df$x)`.
The proportion of cases for which the logical vector is TRUE is `r sum(problem_two_df$vec_logic, na.rm=TRUE) / nrow(problem_two_df)`

Scatterplot of y vs x with color points using the logical variable:
```{r}

ggplot(problem_two_df, aes(x = x, y = y, color = vec_logic)) + geom_point()
```

Scatterplot of y vs x with color points using the numeric variable:
```{r}

ggplot(problem_two_df, aes(x = x, y = y, color = vec_numeric)) + geom_point()
```

Scatterplot of y vs x with color points using the factor variable:
```{r}

ggplot(problem_two_df, aes(x = x, y = y, color = vec_factor)) + geom_point()
```
